Rigidity of Expectations: Additional Evidence from Density Forecasts of Professionals and Households

### 2019-07-06 paper/Draft.tex
font size change to 12pt
clarify the inflation measures. SPF core CPI and core PCE. SCE headline CPI.

\documentclass[]{article}
\usepackage{amsmath}
\usepackage{adjustbox}
\usepackage{graphicx}
\usepackage[a4paper,bindingoffset=0.2in,%
left=1in,right=1in,top=1.2in,bottom=1.2in,%
footskip=.25in]{geometry}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=blue,
	filecolor=blue,      
	urlcolor=blue,
	citecolor=blue
}
\usepackage{natbib}
\usepackage{rotating}
\usepackage{threeparttable}
\usepackage{caption}
\usepackage{subcaption}

@@

	This paper is related to three strands of literature. First, the recent decade has seen an emergence of the empirical work directly testing and evaluating the various theories on expectations theories using survey data. For instance, \citet{mankiw2003disagreement}, \citet{carroll2003macroeconomic}, \citet{branch2004theory}. More recent examples include \citet{coibion2018firms} on firms' managers. In addition to testing particular sets of theories, there is also a number of papers that show people's expectations are driven by idiosyncratic demographics and macroeconomic history experienced(\citet{malmendier2015learning}, \citet{das2017socioeconomic},etc.). In terms of the methodology, this paper is closest to \citet{coibion2012can} and \citet{fuhrer2018intrinsic}.  However, all of these research simply rely upon point forecasts instead of density forecast or surveyed uncertainty. This is one theme on which this paper differs from the existing literature.

	Second, the literature that has been originally developed under the theme of forecast efficiency provides a framework analyzing the dynamics of uncertainty useful for the purpose of this paper. The focus of the forecasting efficiency literature is evaluating forecasters' performance and improving forecasting methodology, but it can be adapted into testing the context of understanding expectation formation of different types of agents. This is especially relevant to this paper as I focus on the uncertainty.

@@

	\begin{figure}[ht]
		\centering
		\includegraphics[width=13cm]{figures/ni_illustration.png}  \\
		{\footnotesize Note: the top two figures plot the nowcasting of $y_{t+k}$  at time $t+k$ and forecasting of $y_{t+h}$ at time $t+k$  based on a Kalman filtering together with FIRE after a one-time shock of one unit at time $t$. The nowcasting and forecasting uncertainty are correspondingly plotted in the bottom. Both private signal and public signal have the conditional variance equal to long-run variance of y, i.e.  $\sigma_\epsilon = \sigma_\xi = \sigma_y$. The plotted signals are are a pair of particular draw from their respective conditional distribution. }
		\caption{Nowcasting and Forecasting in Noisy Information}
		\label{IllustrateNI}
	\end{figure}

@@

	\begin{table}[ht]
		\caption{Information of Data}
		\label{DataInfo}
		\begin{tabular}{lll}

			\hline 
			& SCE & SPF        \\
			\hline 
			Time period                                    & 2013-present                            & 2007-present             \\
			Frequency                                      & Monthly                                 & Quarterly                \\
			Sample Size                                    & 1,300                                   & 30-50                    \\
			Aggregate Var in Density                       & 1-yr  and 3-yr ahead inflation          & 1-yr CPI and PCE         \\
			Individual Var in Density                      & 1-yr earning growth                     & No                       \\
			Panel Structure                               & stay up to 12 months                    & average stay for 5 years \\
			Demographic Info                        & Education, Income, Age, Location        & Industry    \\
			\hline 
		\end{tabular}
	\end{table}

	Throughout the paper, I use two measures of inflation, headline CPI and PCE . The forecasts-based measures are computed and used correspondingly.

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{adjustbox}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage[a4paper,bindingoffset=0.2in,%
left=1in,right=1in,top=1.2in,bottom=1.2in,%
footskip=.25in]{geometry}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=blue,
	filecolor=blue,      
	urlcolor=blue,
	citecolor=blue
}
\usepackage{natbib}
\usepackage{rotating}
\usepackage{threeparttable}
\usepackage{caption}
\usepackage{subcaption}

@@

	This paper is related to three strands of literature. First, the recent decade has seen an emergence of the empirical work directly testing and evaluating the various theories on expectations theories using survey data. For instance, \citet{mankiw2003disagreement}, \citet{carroll2003macroeconomic}, \citet{branch2004theory}. More recent examples include \citet{coibion2018firms} on firms' managers. In addition to testing particular sets of theories, there is also a number of papers that show people's expectations are driven by idiosyncratic demographics, cognitive abilities and macroeconomic histories experienced(\citet{malmendier2015learning}, \citet{das2017socioeconomic} and \citet{d2019iq}, etc.). In terms of the methodology, this paper is closest to \citet{coibion2012can} and \citet{fuhrer2018intrinsic}.  However, all of these research simply rely upon point forecasts instead of density forecast or surveyed uncertainty. This is one theme on which this paper differs from the existing literature.

	Second, the literature that has been originally developed under the theme of forecast efficiency provides a framework analyzing the dynamics of uncertainty useful for the purpose of this paper. The focus of the forecasting efficiency literature is evaluating forecasters' performance and improving forecasting methodology, but it can be adapted to test the theories of expectation formation of different types of agents. This is especially relevant to this paper as I focus on the uncertainty.

@@

	\begin{figure}[ht]
		\centering
		\includegraphics[width=13cm]{figures/ni_illustration.png}  \\
		\begin{flushleft}
		{\footnotesize Note: the top two figures plot the nowcasting of $y_{t+k}$  at time $t+k$ and forecasting of $y_{t+h}$ at time $t+k$  based on a Kalman filtering together with FIRE after a one-time shock of one unit at time $t$. The nowcasting and forecasting uncertainty are correspondingly plotted in the bottom. Both private signal and public signal have the conditional variance equal to long-run variance of y, i.e.  $\sigma_\epsilon = \sigma_\xi = \sigma_y$. The plotted signals are are a pair of particular draw from their respective conditional distribution. }
		\end{flushleft}
		\caption{Nowcasting and Forecasting in Noisy Information}
		\label{IllustrateNI}
	\end{figure}

@@

	\begin{table}[ht]
		\caption{Information of Data}
		\label{DataInfo}
		\begin{tabularx}{\textwidth}{|X|X|X|}
			\hline 
			& SCE & SPF        \\
			\hline 
			Time period                                    & 2013-2019                           &
			 2007-2019             \\
			 \hline 
			Frequency                                      & Monthly                                 & Quarterly                \\
			\hline 
			Sample Size                                    & 1,300                                   & 30-50                    \\
			\hline 
			Var in Density                       & 1-yr and 2-yr-ahead inflation          & 1-yr-ahead GDP deflator, Core CPI and Core PCE         \\
			\hline 
			Panel Structure                               & stay up to 12 months                    & average stay for 5 years \\
			\hline 
			Individual Info                        & Education, Income, Age, Location        & Industry    \\
			\hline 
		\end{tabularx}
	\end{table}

	Throughout the paper, I use three measures of inflation: headline CPI, core CPI and core PCE . Depending on the specific variable of forecast in the survey series, the realization of the corresponding inflation is used to compute moments such as forecast errors. Specifically, SPF has density forecasts for both core CPI and core PCE \footnote{SPF also  has density forecasts for GDP deflator(for GNP prior to 1992:Q1) going back to 1968. Since the ranges of the values as we as the definition have not been consistent over time. I do not use it in this paper.}. For SCE, as the households responders are asked about the overall inflation, It is the most appropriate to be interpreted as  headline CPI inflation. To simplify the expression, from now on, core CPI and core PCE are simply referred to as CPI and PCE, respectively.
